### 2020-11-30 book/bookdown-minimal-master/1_Methods.Rmd
Remove nested structure from emmeans::ref_grid

load("~/Documents/Git/EP1162/data/ep1162_Data.RData")
load("~/Documents/Git/EP1162/data/ep1162_plotData.RData")

@@

  # Diameter class analysis
  figNums("diamVol.plot", caption="Mean plot volume distribution by diameter class and treatment, for two periods.  Black line shows total volume.  Volume calculated on trees meeting 18.5cm DBH limit.  Spruce and fir contribution to volume by diameter class are shown with bar plots.  Douglas-fir volume contribution by diameter class is not shown.",display="FALSE")

@@

### Stand-level volume and basal area estimates
Total and merchantable volumes were estimated for trees using species- and region-specific equations [@nigh2016] based on tree height and DBH.  Consistent with forest management practices in British Columbia, we estimated volumes for trees meeting a minimum diameter utilization limit of 18.5cm DBH.  For trees lacking a tree height measurement, we estimated height by fitting a single diameter-height model developed with a linear mixed effects model, with log-transformed DBH as a fixed effect, and tree nested within plot as a random effect.  Trees with broken tops were omitted from diameter-height models. We summed tree-level estimates of volume and basal area to generate stand-level estimates for each plot. To test the hypothesis that changes in stand attributes over 27 years varied by treatment, we fit separate linear models with stand total volume, basal area, and density as response variables, and the interaction between treatment and year as the independent variable.

>>>>

@@

  # Diameter class analysis
  figNums("diamVol.plot", caption="Mean plot volume distribution by diameter class and treatment, for two periods.  Black line shows total volume.  Volume calculated on trees meeting 17.5cm DBH limit.  Spruce and fir contribution to volume by diameter class are shown with bar plots.  Douglas-fir volume contribution by diameter class is not shown.",display="FALSE")

@@

### Stand-level volume and basal area estimates
Total and merchantable volumes were estimated for trees using species- and region-specific equations [@nigh2016] based on tree height and DBH.  Consistent with forest management practices in British Columbia, we estimated volumes for trees meeting a minimum diameter utilization limit of 17.5cm DBH.  For trees lacking a tree height measurement, we estimated height by fitting a single diameter-height model developed with a linear mixed effects model, with log-transformed DBH as a fixed effect, and tree nested within plot as a random effect.  Trees with broken tops were omitted from diameter-height models. We summed tree-level estimates of volume and basal area to generate stand-level estimates for each plot. To test the hypothesis that changes in stand attributes over 27 years varied by treatment, we fit separate linear models with stand total volume, basal area, and density as response variables, and the interaction between treatment and year as the independent variable.

### 2020-11-30 book/bookdown-minimal-master/2_results.Rmd
Remove nested structure from emmeans::ref_grid

  # create a reference grid for five different diameter classes (converted to BA)
  rg<-ref_grid(fitBA,at=list(Init=d2ba(c(10,20,30,40,50))),type="response")

@@

  # create a reference grid for five different diameter classes (converted to BA)
  rg.pbaipct<-ref_grid(fitBApct,at=list(Init=d2ba(c(10,20,30,40,50))),type="response")

@@

The tree-level basal area increment model indicated that interactions between log-transformed tree size, species and treatment explained `r round(r.squaredGLMM(fitBA)[1]*100,1)`% (marginal R²) of the variation in tree log-transformed 1992-2019 BAI (`r tableNums("fitBA.table",display="cite")`).  Estimated marginal means from the model showed that between species and among treatments, larger trees increased basal area more than smaller trees (`r figNums("treeGrowth.plot",display="cite")`).  Across diameter classes, spruce trees had more basal area increment than fir in the high-removal treatment unit, whereas growth between the two species was close to equal in the low-removal and control units (`r appNums("sppContrast.table",display="cite")`).  Spruce basal area increment in the high-removal unit exceeded spruce growth in both other units (`r appNums("trtContrast.table",display="cite")`).  Spruce growth was similar between the low-removal and control units.  Fir basal area increment in the high- and low-removal units was higher than the control unit, and there were no significant differences between the two harvested units (`r appNums("trtContrast.table",display="cite")`).

@@

  # Remove volumes below 18.5cm DBH
  mutate(vol=replace(vol,DBH<18.5,0)) %>%

@@

# Contrasts
vol.contrast <- contrast(vol.means,interaction="pairwise")
ba.contrast <- contrast(ba.means,interaction=c("pairwise")) 
sph.contrast <- contrast(sph.means,interaction=c("pairwise"))

@@

# Create contrasts table
standFit.contrasts<-
  data.frame(Variable="Volume",vol.contrast) %>% 
  rbind(data.frame(Variable="Basal area",ba.contrast)) %>% 
  rbind(data.frame(Variable="Stand density",ba.contrast)) %>% 
  filter(Year_pairwise %in% c("1992 - 2019","1997 - 2019")) %>% 
  dplyr::select(Variable,`Year contrast`="Year_pairwise",`Treatment contrast`="BA.Target_pairwise",everything()) %>% 
  rename(`t ratio`="t.ratio",
         `p value`="p.value") %>% 
  mutate_at(vars(estimate:`t ratio`),round,3) %>% 
  mutate_at(vars(`p value`),round,5)

figNums("standFit.X",display=FALSE)

@@

Plot means of volume and basal area show that from 2009-2019, stand volume and basal area converged between both harvested treatment units, and were less than the control (
`r figNums("standPlots",display="cite")`).  The slope of all attributes showed that the low RBA accrued volume, density and diameter faster than the other units over the 27-year period of this study.  A linear model

@@

@@

@@

```{r}
diamVol.plot

@@

*Note that in this figure, we show volume for all diameter classes.  This is different than the stand-level estimate of volume, where we only estimate volume for trees that would be considered merchantable under conventional harvest in British Columbia (i.e., minimum 18.5cm DBH).*

@@

# create stand-level plots by diameter distribution
sppVol.plot<-sppComp.plot("Vol")

>>>>

  # create a reference grid for five different diameter classes (converted to BA)
  rg<-ref_grid(fitBA,at=list(Init=d2ba(c(10,20,30,40,50))),type="response",nesting=NULL)

@@

  # create a reference grid for five different diameter classes (converted to BA)
  rg.pbaipct<-ref_grid(fitBApct,at=list(Init=d2ba(c(10,20,30,40,50))),type="response",nesting=NULL)

@@

The tree-level basal area increment model indicated that interactions between log-transformed tree size, species and treatment explained `r round(r.squaredGLMM(fitBA)[1]*100,1)`% (marginal R²) of the variation in tree log-transformed 1992-2019 BAI (`r tableNums("fitBA.table",display="cite")`).  Estimated marginal means from the model showed that between species and among treatments, larger trees increased basal area more than smaller trees (`r figNums("treeGrowth.plot",display="cite")`).  Species contrasts showed that spruce trees increased  basal area more than than fir in the high-removal treatment unit, whereas basal growth between the two species was close to equal in the low-removal and control units (`r appNums("sppContrast.table",display="cite")`).  Within each species, treatment contrasts showed that spruce basal area increment in the high-removal unit exceeded spruce growth in both other units (`r appNums("trtContrast.table",display="cite")`), whereas fir basal area increment in the high- and low-removal units was higher than the control unit, and there were no significant differences between the two harvested units (`r appNums("trtContrast.table",display="cite")`).

@@

  # Remove volumes below 17.5cm DBH
  mutate(vol=replace(vol,DBH<17.5,0)) %>%

@@

# Slope Contrasts
vol.contrast <- contrast(vol.means,interaction="pairwise")
ba.contrast <- contrast(ba.means,interaction=c("pairwise")) 
sph.contrast <- contrast(sph.means,interaction=c("pairwise"))

# Mean contrasts
volM.contrast <- pairs(vol.means,by="Year")
baM.contrast <- pairs(ba.means,by="Year")
sphM.contrast <- pairs(sph.means,by="Year")

@@

# Create slope contrasts table
standFit.contrasts<-
  data.frame(Variable="Volume",vol.contrast) %>% 
  rbind(data.frame(Variable="Basal area",ba.contrast)) %>% 
  rbind(data.frame(Variable="Stand density",sph.contrast)) %>% 
  filter(Year_pairwise %in% c("1992 - 2019")) %>% 
  dplyr::select(Variable,`Year contrast`="Year_pairwise",`Treatment contrast`="BA.Target_pairwise",everything()) %>% 
  rename(`t ratio`="t.ratio",
         `p value`="p.value") %>% 
  mutate_at(vars(estimate:`t ratio`),round,3) %>% 
  mutate_at(vars(`p value`),round,5)

# Create mean contrasts table
standFitM.contrasts<-
  data.frame(Variable="Volume",volM.contrast) %>% 
  rbind(data.frame(Variable="Basal area",baM.contrast)) %>% 
  rbind(data.frame(Variable="Stand density",sphM.contrast)) %>% 
  filter(Year %in% c(1992,2019))

figNums("standFit.X",display=FALSE)
appNums("standFit.contrasts",display=FALSE)

@@

_Observations:_

1. All units increased volume, basal area and stand density over the 1992-2019 period (`r figNums("standPlots",display="cite")`).

2. The rate of change differed between units.  The low RBA unit accrued basal area faster than the other two units over the 1992-2019 period, and this difference was significant (`r appNums("standFit.contrasts",display="cite")`).  The estimated rate of basal area increment in the low RBA unit was higher by by 8.4 and 7.8m² per hectare more than the high RBA and control units, respectively.

3. The low RBA unit also significantly increased its stand density by an estimated 340 stems per hectare more than the control unit. The rate of change in tree density did not differ significantly between the two harvested units.

4. The rate of change in volume in trees over 17.5cm DBH was barely significant between the two harvested units, and there were no significant differences between the control and harvested units.

@@

`r appNums("standFit.contrasts",caption="Contrasts of 1992-2019 stand structure development over 1992-2019 between treatments.  Contrasts are shown for three variables: volume, basal area and stand density.")`

@@

standFit.contrasts %>% 
  dplyr::select(-`Year contrast`) %>% 
  flextable() %>% 
   align(align="left",part="all") %>% 
  merge_v(j=c("Variable")) %>% 
  valign(valign="top",part="body") %>% 
    hline(i=c(3,6,9),border=fp_border(color="black")) %>% 
    fix_border_issues() %>% 
  autofit()

@@

```{r}
diamBA.plot

@@

*Note that in this figure, we show volume for all diameter classes.  This is different than the stand-level estimate of volume, where we only estimate volume for trees that would be considered merchantable under conventional harvest in British Columbia (i.e., minimum 17.5cm DBH).*

@@

# create stand-level plots by diameter distribution
sppVol.plot<-sppComp.plot("BA")
